
### SQL and APIs project


A client, Acme Corporation, has a PostgreSQL database containing basic details of their `employees`, the `teams` they belong to and `feedbacks` they have given on their employment (text snippets detailing positive and negative experiences, suggestions for improvement etc).

Acme have hired us to first do some basic analysis of the data they store, and then hopefully perform a quantitative analysis of the emotional state of their corporation based on the stored `feedbacks`.



# MVP - SQL

Create a local PostgreSQL database called `acme_employees` and seed it using the SQL commands stored in file `acme_employees.sql`.

In `Terminal`: 

```bash
createdb acme_employees
psql -d acme_employees -f acme_employees.sql
```

Take some time to explore your new `acme_employees` database using the `psql` command line interface (remember the `\dt` and `\d+` commands). Look for tables and the relationships between tables. How many records are there in each table? Identify any fields corresponding to primary keys and foreign keys in each table?

Three tables: `employees` (200 records), `feedbacks` (60 records) and `teams` (11 records)
PK is 'primary key', FK is 'foreign key':

-  `employees`: PK `id`, FK `team_id` NOT NULL, referenced by `feedbacks`.
-  `teams`: PK `id`, no FK, referenced by `employees`
-  `feedbacks`: PK `id`, FK `employee_id` NOT NULL

Relationships:

- One team has many employees
- One employee has many feedbacks

Create a connection from R to your local `acme_employees` database.




```{r}
library(keyring)
library(RPostgreSQL)

keyring_unlock("local")
username <- key_get(service = "postgresql_username", keyring = "local")
password <- key_get(service = "postgresql_password", keyring = "local")
keyring_lock("local")

db_connection <- dbConnect(drv = PostgreSQL(), user = username, password = password, dbname = "acme_employees", host = "localhost")

rm(username, password)
```

Create a new `data.frame` containing the details of all `employees` held in the database

```{r}
all_employees <- dbGetQuery(db_connection,
           "SELECT * FROM employees")
all_employees
```

Retrieve the records for all `employees` again, but this time add a field `full_name` containing the `first_name` and `last_name` of each employee separated by a space.

```{r}
dbGetQuery(db_connection,
           "SELECT *, CONCAT(first_name, ' ', last_name) AS full_name 
           FROM employees")
```

Retrieve the details of all `employees` again, but this time listing their team `name` (as `team_name`), rather than their `team_id`.

```{r}
dbGetQuery(db_connection,
           "SELECT e.first_name, e.last_name, e.email, t.name AS team_name
           FROM employees AS e LEFT JOIN teams AS t
           ON e.team_id = t.id")
```

Do a breakdown of the number of `employees` per team (here we want two columns: `team_name` and `num_employees`). Order the table in decreasing number of employees. Include all `teams`, even if they have no employees.

```{r}
dbGetQuery(db_connection,
           "SELECT t.name AS team_name, COUNT(e.id) AS num_employees
           FROM employees AS e RIGHT JOIN teams AS t
           ON e.team_id = t.id
           GROUP BY t.name
           ORDER BY num_employees DESC")
```

AcmeCorp management want to send an email survey to each of their `employees`. However, they are worried that some of the `employees` may not receive the email either because their email addresses may be invalid (in that they don't contain an '@' symbol), or they may not have a stored email address at all. Can you find all these `employees`?

```{r}
dbGetQuery(db_connection,
           "SELECT * FROM employees
           WHERE email NOT LIKE '%@%' OR email IS NULL")
```

You are chatting with one of the AcmeCorp employees, and she mentions that one of the `teams` in the corporation has some `employees` whose `first_name` contains only their first initial, e.g. "D" instead of "Dionne", "C" instead of "Charles" (they had a really formal administrator a few years back). Can you find the `name` of the team that she is most likely referring to (i.e. which team has the highest number of employees with single initial `first_names`)`?

```{r}
dbGetQuery(db_connection,
           "SELECT COUNT(*) AS num_only_first_initial, t.name AS team_name 
           FROM employees AS e LEFT JOIN teams AS t
           ON e.team_id = t.id
           WHERE e.first_name LIKE '_'
           GROUP BY team_name
           ORDER BY num_only_first_initial DESC NULLS LAST
           LIMIT 1")
```


# MVP - APIs

AcmeCorp want a way to assess the emotional content of the `feedbacks` left by their employees. In a previous project you heard about the `https://apiv2.indico.io/emotion` endpoint of the Indico APIs suite, and you know that this might help solve the problem. You `POST` a textual message to the endpoint, and it responds with a measure of the emotions in the message. For example

```{r}
message <- "I love writing R code!"
# POST to API and get content() of response
# will be a list with structure
response_list <- list(
  results = list(
    anger = 0.05099975, 
    fear = 0.04647598, 
    joy = 0.68175, 
    sadness = 0.0226995, 
    surprise = 0.1980747
  )
)
```

If you haven't yet got an API key for the Indico API suite, apply for one now (you'll need to set up a **free** Indico account, your API key will be on your account homepage), receive it, and then store it in an appropriately named key in your `remote` keyring.

```{r, eval = FALSE}
keyring_unlock("remote")
key_set("indico_api_key")
keyring_lock("remote")
```

Write a `get_emotions_from_indico_api()` function that takes in `api_key`, `url` and `message`, `POST`s the `message` to the Indico `emotions` endpoint, and returns the `content(response)`. Try your function out one-by-one on the following two trial messages, and investigate the structure of the response contents:

```{r, eval = FALSE}
messages <- c("I really like working here, great place, decent perks", "Hmm, not so sure I'll stay: long working hours and low pay, bad morale")
```

(1) This function will contain a `POST` request, but you can pass query params just as you did for `GET` requests previously. (2) The params to be passed look like `params <- list(api_key = api_key, data = message)`, (3) don't forget to get your API key from your keyring. This should be passed in to your function, and then into the `params` list.

```{r}
library(httr)
get_emotions_from_indico_api <- function(api_key, url, message){

  params <- list(
    api_key = api_key,
    data = message
  )

  response <- POST(url, query = params)

  return(content(response))
}
```

```{r, eval=FALSE}
messages <- c(
  "I really like working here, great place, decent perks", 
  "Hmm, not so sure I'll stay: long working hours and low pay, bad morale"
)

keyring_unlock("remote")
indico_api_key <- key_get("indico_api_key")
keyring_lock("remote")
indico_url <- 'https://apiv2.indico.io/emotion'

from_message_1 <- get_emotions_from_indico_api(indico_api_key, indico_url, messages[1])
from_message_1
from_message_2 <- get_emotions_from_indico_api(indico_api_key, indico_url, messages[2])
from_message_2
str(from_message_1)
```


# Combining SQL and APIs

AcmeCorp would like mean values of each of the five emotions, averaging over all of the employee messages they have stored in their database. Based on the work you did above, pass all of the messages held in the `feedbacks` table through the `emotions` API endpoint, storing the output in a data frame. Next, calculate the mean value of each of the emotions stored in this data frame.  

```{r, eval=FALSE}
# get all feedbacks
messages <- dbGetQuery(db_connection, "SELECT message FROM feedbacks")
messages <- messages$message

# pass all through emotions endpoint and store results
results <- map(messages, get_emotions_from_indico_api, api_key = indico_api_key, url = indico_url)
flatten_results <- map(results, flatten)
results <- map_dfr(flatten_results, as.list)

# now get mean of each emotion
map_df(results, mean, na.rm = TRUE)
```

Which team has the most joyful employee? Which, the most angry? 

```{r, eval=FALSE}
# first, get data frame with employee names, team name and message
employees_teams_messages <- dbGetQuery(db_connection, 
           "SELECT e.id, e.first_name, e.last_name, t.name AS team_name, f.message
           FROM employees AS e LEFT JOIN teams AS t
           ON e.team_id = t.id
           INNER JOIN feedbacks AS f
           ON f.employee_id = e.id")

# now fetch all emotions results
messages <- employees_teams_messages$message
results <- map(messages, get_emotions_from_indico_api, api_key = indico_api_key, url = indico_url)
flatten_results <- map(results, flatten)
results <- map_dfr(flatten_results, as.list)

employees_teams_messages_emotions <- cbind(employees_teams_messages, results)

# team with most joyful employee
subset(employees_teams_messages_emotions, joy == max(joy, na.rm = TRUE))$team
# team with most angry employee
subset(employees_teams_messages_emotions, anger == max(anger, na.rm = TRUE))$team
```


